\documentclass{llncs}       % onecolumn (second format)

\usepackage{listings}
\usepackage[longend,ruled]{algorithm}
\usepackage{algorithmic}
\usepackage{fancyvrb}
\usepackage{graphicx}
\usepackage{tikz}
%\usepackage{refs,xspace}
%\usepackage{macros}
\usepackage{times}
%\usepackage{color,graphicx,enumerate}
%\usepackage{amsmath,amssymb,amsfonts,xspace}
\usepackage{url}
%\usepackage[T1]{fontenc}
\usepackage{floatflt}
\usepackage{epsfig,pifont}

\usepackage[square,sort&compress]{natbib}
\usepackage{mdframed}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{amssymb,latexsym,amsfonts,amsmath}
\usepackage{wrapfig}
\usepackage{hyperref}
\usepackage{paralist} 
\usepackage{booktabs}
\usepackage{multirow}
%\usepackage{smallsec}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsExtensions{.pdf}
%\renewcommand{\baselinestretch}{0.95} 
\usepackage[ruled,shortend,algo2e]{algorithm2e}
\usepackage{comment}

\definecolor{mygray}{rgb}{0.95,0.95,0.95}
\definecolor{darkgreen}{rgb}{0,.45,0}
\lstset{language=C}
\lstset{backgroundcolor=\color{mygray}}
\lstset{basicstyle=\ttfamily\small}
\lstset{numbers=left, numberstyle=\small, stepnumber=1, numbersep=10pt}
\lstset{keywordstyle=\color{red}\emph} %\bfseries
\lstset{commentstyle=\color{darkgreen}}
\lstset{numberbychapter=false}
% these are the syntax-highlighting color definitions we will use.
\lstdefinestyle{customc}{
  belowcaptionskip=1\baselineskip,
  breaklines=true,
  frame=single,
  language=C,
  captionpos=b,
  otherkeywords={uint8_t, assert, assume, bool, havoc, sensitive},
  showstringspaces=false,
  numbers=left,
  basicstyle=\ttfamily,
  keywordstyle=\bfseries\color{green!40!black},
  commentstyle=\color{purple!40!black},
  identifierstyle=\color{blue},
  stringstyle=\color{orange},
}
\lstset{escapechar=@,style=customc}

\newcommand{\code}[1]{$\mathtt{#1}$}
\newcommand{\mcode}[1]{\mathtt{#1}}

\usetikzlibrary{arrows,backgrounds,decorations,decorations.pathmorphing,positioning,fit,automata,shapes.multipart,shapes,snakes}
\usetikzlibrary{patterns,matrix}
\usepackage[normalem]{ulem}
%\newenvironment{Example}{\begin{example}}{\hfill$\blacksquare$\end{example}}
\setlength{\abovecaptionskip}{2pt}
\addtolength{\textfloatsep}{-1ex}
%\usepackage{flushend}
% To eliminate orphans & widows
\clubpenalty = 10000
\widowpenalty = 10000
\displaywidowpenalty = 10000

\begin{document}

\include{macros}
\title{Lazy Self-Composition for Security Verification}

%\author{Blank for double-blind review} 
\author{ Weikun Yang \and 
         Pramod Subramanyan \and 
         Yakir Vizel \and
         Aarti Gupta \and 
         Sharad Malik} 

\institute{Princeton University}

\maketitle

\begin{abstract}
\begin{comment}
We present a novel path-sensitive information flow analysis technique.

Secure information flow analysis is meant to detect if high-security assets can leak during an execution of a program. One approach for performing a precise information flow analysis, is by a reduction to safety verification by means of \emph{self-composition}. However, this reduction is intractable since it requires two copies of the program to be created, on which a safety property is checked. A second, less precise approach, is by using \emph{taint analysis}. Taint analysis provides a sound approximation for secure information flow analysis. While more efficient than self-composition, this approach often results in many ``false'' alarms, i.e. reporting leaks that do not exist.

In order to bridge the gap between the above two approaches, in this paper we present \Ifc, a novel, precise and efficient, secure information flow analysis technique. \Ifc relies on an interplay between a symbolic path-sensitive taint analysis and self-composition, where taint analysis guides self-composition and vice-versa. As a result of this interplay, full self-composition is never required, yet, the overall technique is sound and complete.

We implemented a prototype of \Ifc on top of SeaHorn and evaluated it on challenging examples. The experimental results show the potential of our approach.


%% aarti's suggestion below

We present a novel algorithm for verifying secure information flow that guarantees the absence of information leaks from high-security inputs to low-security outputs in a program. A classic approach for verifying secure information flow is to reduce the problem to safety verification on a \emph{self-composition}, where two copies of the program are created~\cite{BartheCsfw04}.
Although this reduction is sound and complete, it is challenging to perform safety verification in practice, especially on two copies of a program.
Another, less precise approach, is by using \emph{taint analysis} based on an information flow type system. Taint analysis provides a sound approximation for secure information flow. While more scalable than the self-composition approach, it often results in many false alarms, i.e., reporting leaks that do not exist.

To bridge the gap between these two approaches, in this paper we present \Ifc, a novel, precise, and efficient verifier for secure information flow. \Ifc relies on an interplay between a symbolic path-sensitive taint analysis and self-composition, where taint analysis guides self-composition and vice-versa. As a result of this interplay, full self-composition is never required, yet, the overall technique is sound and complete.
We implemented a prototype of \Ifc on top of SeaHorn~\cite{Seahorn} and evaluated it on challenging examples. The experimental results show the potential of our approach.
\end{comment}

%%%%%%%%%%%%%

The secure information flow problem has many applications in checking security properties in programs, e.g., that there are no information leaks from high-security inputs to low-security outputs. 
In this paper, we present \Ifc, a novel, precise, and efficient verifier for secure information flow. \Ifc improves the classical  self-composition approach (due to Barthe et al.) by making it lazier, in contrast to an eager upfront translation into two copies of a program. This lazy duplication is guided by taint analysis.
We implemented a prototype of \Ifc on top of SeaHorn tool and evaluated it on challenging examples. The experimental results show the potential of our approach.

\end{abstract}

\section{Introduction}
\label{sec:intro}
\input{intro}

\section{Motivating Examples}
\label{sec:motiv}
\input{motiv}

\section{Preliminaries}
\label{sec:prelim}
\input{prelim}

%\section{Lazy Self-Composition}
\label{sec:algo}
\input{algo}

\section{Implementation and Experiments}
\label{sec:exp}
\input{exp}

\section{Related Work}
\label{sec:related}
\input{related}

\section{Conclusions and Future Directions}
\label{sec:concl}
\input{concl}

\bibliographystyle{splncs03}  

\bibliography{b1}

\end{document}
